% Enterprise JSF project.
%
% Copyright 2022 e-Contract.be BV. All rights reserved.
% e-Contract.be BV proprietary/confidential. Use is subject to license terms.

\chapter*{References}
\addcontentsline{toc}{chapter}{References}
Within this section, we give an overview of books in chronological order that contain some chapters on custom JSF component development.


\section*{Core JavaServer Faces, Third Edition}
This book \cite{geary2010core} from 2010 by David Geary and Cay Horstmann covers JSF version 2.0
Chapter 9 covers composite components.
Chapter 11 covers custom components.


\section*{Pro JSF and HTML5 - Second Edition}
The book Pro JSF and HTML5 - Second Edition \cite{wadia2014pro} published in 2014 covers JSF version 2.2 and comes closest to what we are looking for.
Besides a very good in-depth overview of JSF, there is a lot of material on how to construct custom JSF components.


\section*{Mastering JavaServer Faces 2.2}
The book Mastering JavaServer Faces 2.2 \cite{leonard2014mastering} by Anghel Leonard, published in 2014, conversion JSF version 2.2.
Chapter 10 covers development of custom components.
 
 
 \section*{PrimeFaces Cookbook - Second Edition}
 This book \cite{10.5555/2834507} covers PrimeFaces.
 Unfortunately a chapter on writing custom components, which was available within the first edition of this book, did not make it within the second edition.
 This chapter however, is still available for download \cite{PrimeFacesCookbookChapter11}.


\section*{The Definitive Guide to JSF in Java EE 8}
This book \cite{Scholtz2018TheDG} from 2018, written by the grand masters Bauke Scholtz and Arjan Tijms, covers JSF version 2.3.
There is chapter 11 that covers custom components.


\section*{Practical JSF in Java EE 8}
This book \cite{muller2018practicalJSF} from 2018, written by Michael Müller, convers JSF version 2.3.
Chapter 27 touches on custom composite components.



